Species motif participation provides unique information about species risk of extinction

### 2023-04-18 manuscript/JAE_reply.tex
Updates from Overleaf

@@

                \textbf{R:} Probably the Reviewer is confused... I am confused by the comment.

>>>>

                \textbf{R:}
            The sentence the reviewer refer to is "How disturbances such as population decline or extinction lead to the loss of species via  direct and indirect effects (i.e., secondary extinctions) has long been a vibrant area of research (Curtsdotter et al., 2011a; Dunne & Williams, 2009; Ekl ̈of & Ebenman, 2006 Santos et al., 2021) " Here

@@

                \textbf{R:} Can the Reviewer provide a reference for their claim that apparent competition is destabilising? Is this a very general trend, or something that only happens in rare cases? Note that we do not claim that apparent competition is \emph{never} destabilising, only that more frequent participation in the apparent competition motif is associated with greater persistence on average. Moreover, we cannot think of a reason why a bottom-up disturbance (i.e., loss of some resources) would be destabilising in an apparent competition motif - surely it is more likely that a predator with two prey will persist after losing one than than two predator sharing one prey will persist if that prey is lost. We can see how it is possible for strong \emph{top-down} effects to destabilise an apparent competition motif (e.g., if a predator suddenly exerts much more pressure on one of the prey than the other), but this paper is concerned with bottom-up effects only. Given the earlier work showing that Bayesian networks capture the majority of secondary extinctions that occur in topological models that include both top-down and bottom-up effects, we can conclude that top-down effects are much less common as a cause of extinction than bottom-up effects. Of course they can still be equally interesting and may be strong in particular focal systems, but as our simulations do not reflect a specific community we have no reason to expect top-down effects to have a stronger role than in the networks used to validate the Bayesian network approach.

### 2023-04-18 manuscript/supplement_motifs.tex
Updates from Overleaf

\section{Participation and network properties} [not updated yet]

    Motif participation varies with global-scale network properies.
    % Updated with network-level random effect
    \begin{table}[hb!]
        \centering
        \caption{Motif participation varies with size and connectance. Here we give the range and mean ($\pm$ standard deviation) proportion of each motif in species' motif participation vectors. 
        We also show the coefficients ($\beta$) for network size, connectance, and their interaction in general linear mixed-effect models predicting the proportion of each motif in a participation vector.
        These models also included a random effect of network ID, as species within the same network are not independent and were fit with a binomial error distribution and logit link function.
        Predictors were not centered or standardized before model fitting. Motifs are ordered according to mean proportion.}
        % Relevant lms are pchainlm, pomnilm, papplm, pdirlm}
        \label{tab:partic_vs_SC}   
        \footnotesize
        \begin{tabular}{c|c c c | c c | c c | c c}
             &  &  &  & \multicolumn{2}{c|}{Size} & \multicolumn{2}{c|}{Connectance} & \multicolumn{2}{c}{Interaction} \\
            Motif & Min & Max & Mean (SD) & $\beta$ & $p$ & $\beta$ & $p$ & $\beta$ & $p$ \\
            \hline
            3-sp Chain & 0 & 1 & 0.249 (0.127) & 2.83$\times10^{-4}$ & \textless0.001 & -0.0982 & 0.053 & -3.66$\times10^{-3}$ & \textless0.001 \\
            App. Comp & 0.250 & 0.765 & 0.422 (0.0645) & -5.85$\times10^{-4}$ & \textless0.001 & -0.840 & \textless0.001 & 2.69$\times10^{-3}$ & 0.003 \\
            Dir. Comp & 0.0501 & 0.359 & 0.179 (0.0407) & -6.99$\times10^{-5}$ & 0.450 & -0.636 & \textless0.001 & 6.46$\times10^{-3}$ & \textless0.001 \\
            Omnivory & 0.006 & 0.346 & 0.164 (0.0790) & 3.70$\times10^{-4}$ & \textless0.001 & 1.57 & \textless0.001 & -5.48$\times10^{-3}$ & \textless0.001\\   
            \hline
            \end{tabular}
            \end{table}

    % Updated with network-level random effect
    \begin{figure}[ht!]
        \centering
        \includegraphics[height=.4\textheight]{figures/participation_vs_SC.eps}
        \caption{The proportion of each motif in a consumer's participation vector varied sigificantly with connectance but not with network size or the interaction between size and connectance. Because the interaction was not significant, we show relationships for intermediate values of size and connectance when plotting against the other variable. These relationships were steepest for apparent competition and omnivory.}
        \label{fig:roles_vs_SC}
    \end{figure}

    Motif participation also varied with a species' degree and trophic level.
    \begin{table}[hb!]
        \centering
        \caption{Motif participation varies with in-degree and trophic level. Here we show the coefficients for models relating the frequency of each motif in a species' participation vector to in-degree or trophic level (8 models in total). All relationships were significant ($\alpha$=0.05).
        All models were fit with binomial error distribution and logit link function.
        Predictors were not centered or standardized before model fitting. Motifs are ordered according to mean proportion.}
        % Relevant lms are pchainlm, pomnilm, papplm, pdirlm}
        \label{tab:partic_vs_SC}   
        \footnotesize
        \begin{tabular}{c|c c }
            Motif & in-degree & TL \\
            \hline
                3-sp Chain & \textbf{-6.99$\times10^{-3}$} & \textbf{9.84$\times10^{-2}$} \\
            Omnivory & \textbf{2.29$\times10^{-2}$} & \textbf{-2.27$\times10^{-1}$}\\
            App. Comp & \textbf{-8.76$\times10^{-3}$} & \textbf{2.40$\times10^{-1}$}\\
            Dir. Comp & \textbf{-1.43$\times10^{-3}$} & \textbf{-2.52$\times10^{-1}$}\\
            \hline
            \end{tabular}
            \end{table}

>>>>

\section{Participation and network properties} [updated]

    Motif participation varies with global-scale network properies.
    % Updated with network-level random effect
    \begin{table}[hb!]
        \centering
        \caption{Motif participation varies with size and connectance. Here we give the range and mean proportion of each motif in species' motif participation vectors. 
        We also show the coefficients ($\beta$) for network size (S), connectance (C), and their interaction in general linear models predicting the proportion of each motif in a participation vector.
        Coefficients in \textbf{bold} are significant ($\alpha$=0.05).
        We also show the coefficients for similar models relating the proportion of each motif to in-degree or trophic level.
        All models were fit with a binomial error distribution and logit link function.
        Predictors were not centered or standardized before model fitting. Motifs are ordered according to mean proportion.}
        % Relevant lms are pchainlm, pomnilm, papplm, pdirlm}
        \label{tab:partic_vs_SC}   
        \footnotesize
        \begin{tabular}{c|c c c | c c c | c | c}
            Motif & Min & Max & Mean & S & C & Int. & in-Deg & TL \\
            \hline
            3-sp Chain & 0 & 1 & 0.249 & \textbf{1.20$\times10^{-3}$} &  -0.652 & \textbf{-1.77$\times10^{-2}$} & \textbf{-6.99$\times10^{-3}$} & \textbf{9.84$\times10^{-2}$} \\
            App. Comp & 0.250 & 0.765 & 0.422 & \textbf{-1.82$\times10^{-3}$} & \textbf{-3.18} & 6.92$\times10^{-3}$ & \textbf{-8.76$\times10^{-3}$} & \textbf{2.40$\times10^{-1}$}\\
            Dir. Comp & 0.0501 & 0.359 & 0.179 & -1.18$\times10^{-3}$ & \textbf{-4.50} & \textbf{4.79$\times10^{-2}$} & \textbf{-1.43$\times10^{-3}$} & \textbf{-2.52$\times10^{-1}$} \\
            Omnivory & 0.006 & 0.346 & 0.164 & \textbf{3.64$\times10^{-3}$} & \textbf{12.5} & \textbf{-4.19$\times10^{-2}$} & \textbf{2.29$\times10^{-2}$} & \textbf{-2.27$\times10^{-1}$}\\   
            \hline
            \end{tabular}
            \end{table}

    \begin{figure}[ht!]
        \centering
        \includegraphics[width=.9\textwidth]{figures/participation_vs_SC.eps}
        \caption{The proportion of each motif in a consumer's participation vector varied sigificantly with connectance but not with network size or the interaction between size and connectance. Because the interaction was not significant, we show relationships for intermediate values of size and connectance when plotting against the other variable. These relationships were steepest for apparent competition and omnivory.}
        \label{fig:roles_vs_SC}
    \end{figure}
